---
title: "Covid19 vaccine study Final2021"
author: "Desiree De La Rosa"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
oldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- oldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
DrugUser = filter(FinalData, DrugUser == "yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Guys

### Barplot

```{r}
dd2 <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>%
mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

# now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")

```

The barchart is showing that 337 people were infected with covid19 and almost 58% were infected when they got the drug. About 42% who were infected received the placebo. It seems that the guys who got the vaccine are one and a half time more likey to get covid, the vaccine appears to be ineffective.

In the first plot on the left, the numbers going up the left side represent the y-axis or the number of people who got covid. The bottom represents the x-axis or whether people got covid or not. As opposed to the second plot which shows the percentage on the y-xis instead. One shows the number of each group who got covid and who didn't with both the vaccine and the placebo and the other shows the same information but in percentage.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that among the guys 58% got covid after taking then vaccine, while about 42% of the guys who took placebo got covid. This shows the vaccine is not effective in preventing covid in males and is not recommended. The guys who took the placebo were less likely to get covid. The guys who gto the vaccine odds of getting covid are 3:2, thats one and a half times more likely than those who took the placebo. 

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the inferential results we can reject the null hypothesis in favor of the alternative, since the p-value is 0.0005801, which is a very tiny number. The null predicted that the deaths related to covid are independent of sex but clearly sex is a factor in the death rates for covid for guys. 

# Gals

### Barplot

```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>%
mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

# now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")

```

Based on this barplot it seems that less gals who got the vaccine than those who took the placebo, which is the opposite of the guys. About 33% of the gals that got the vaccine got covid, meaning the other 67% percent who got covid took placebo. The odds of getting covid for the gals with the vaccine is 1:2, half as likely to get covid than those who took the placebo.

In the first plot on the left, the numbers going up the left side represent the y-axis or the number of people who got covid. The bottom represents the x-axis or whether people got covid or not. As opposed to the second plot which shows the percentage on the y-xis instead. One shows the number of each group who got covid and who didn't with both the vaccine and the placebo and the other shows the same information but in percentage.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that the percentage of gals who took the vaccine and got covid is 33% as we estimated in the barplot. 67% of the gals who took placebo got covid, this shows that the vaccine is effecive in protecting gals against covid. The odds for gals getting covid with the vaccine is 1:2, they are half as likely to get covid than those taking the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the results in the fishers test the p-value is < 2.2e-16 which is very small and this means the results aren't likely to occur by chance so the null hypothesis can be rejected in favor of the alternative.

# Gay Sexual Orientation

### Barplot

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>%
mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

# now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")

```

In this boxplot it appears that about 71% of gay people who got the vaccine still got covid and 29% of them who took the placebo got covid. The odds of getting covid with the vaccine is 2.5:1. People who identified as gay and took the vaccine were three times aas likely to get covis as those who took the placebo.

In the first plot on the left, the numbers going up the left side represent the y-axis or the number of people who got covid. The bottom represents the x-axis or whether people got covid or not. As opposed to the second plot which shows the percentage on the y-xis instead. One shows the number of each group who got covid and who didn't with both the vaccine and the placebo and the other shows the same information but in percentage.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The 71% of the gay participants who took the vaccine got covid, only 29% of those who took the placebo got covid. Based on this information it seems that the vaccine is not effective for the gay community and is not advised. The odds of gettign covid with the vaccine are 2.5:1, meaning gay participants were three times more likely to get covid than those who took the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value in these inferential results is 5.741e-06, which is a very small number as well meaning these results are unlikely to occur by chance therefore we can reject the null hypothesis in favor of the alternative.

# Drug Users

### Barplot

```{r}
dd2 <- DrugUser %>% group_by(infected,treatment) %>% summarize(count=n()) %>%
mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

# now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")

```

Thia barplot shows that about 34% of the people who are drug users and got the vaccine got covid. This means about 66% of the drug using participants who took the placebo got covid. The odds of getting covid with the vaccine are 1:2, meaning those who took the vaccine were half as likely to get covid than thos who took the placebo. 

In the first plot on the left, the numbers going up the left side represent the y-axis or the number of people who got covid. The bottom represents the x-axis or whether people got covid or not. As opposed to the second plot which shows the percentage on the y-xis instead. One shows the number of each group who got covid and who didn't with both the vaccine and the placebo and the other shows the same information but in percentage.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that the percentage of drug used to got the vaccine and still got covid was 33%, 67% of those who took the placebo got covid. It is clear the the vaccine is effective in preventing drug users from getting covid. Getting the vaccine is recommended for this group. The odds of drug users who got the vaccine to get covid were 1:2, meaning they are half as likely to get covid than those who took the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results show that the p-value is 0.002273, this means the results are unlikely to occur by chance. The fishers results show that odds ratio is 0.54 of drug users are half as like to get covid than those who took the placebo. The null hypothesis can be rejected in favor of the alternative since the drug is effective in preventing covid in drug users. 

# Conclusion

The Covid-19 vaccine was not effective in participants who are men or gay.The odds for gay people who took the vaccine were at most at 3:1, thats three times more likely to get covid with the vaccine than with placebo. The Covid-19 vaccine effectively protected drug users, their odds were 1:2 with the vaccine. That's half as likely to get covid than those taking placebo. The vaccine is not reccomended for people who are gay. It is recommended fro individuals who are female or drug users since it was the most effective in those two groups.

